---
title: "Lab2: Intro to Shiny"
author: "Nick McManus"
date: "2023-01-19"
output: html_document
---

```{r setup, include=FALSE, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(shiny)
library(tidyverse)
library(AICcmodavg)
library(equatiomatic)
library(palmerpenguins)
```


## Predicting penguin mass

We're going do to this based on four different characteristics, then check how well these models predict the penguin mass.

```{r}
penguins_clean <- penguins %>% 
  drop_na() %>%  #get rid of rows with NAs in them
  rename(mass = body_mass_g,
         bill_l = bill_length_mm,
         bill_d = bill_depth_mm,
         flip_l = flipper_length_mm)

mdl1 <- lm(mass ~ bill_l + bill_d + flip_l + species + sex + island,
           data = penguins_clean)
```


you can assign formulas to a variable, then plug it into models for ease
```{r}
f1 <- mass ~ bill_l + bill_d + flip_l + species + sex + island
mdl1 <- lm(f1, data = penguins_clean) #same as above

# dropping islands from model bc don't appear statistically significant
f2 <- mass ~ bill_l + bill_d + flip_l + species + sex
mdl2 <- lm(f2, data = penguins_clean)

# now drop bill length bc not as significant
f3 <- mass ~ bill_d + flip_l + species + sex
mdl3 <- lm(f3, data = penguins_clean)

# compare all the models
AIC(mdl1, mdl2, mdl3)
BIC(mdl1, mdl2, mdl3)

# look at corrected AIC
AICcmodavg::AICc(mdl1)  #very similar to AIC bc lots of obs

aictab(list(mdl1, mdl2, mdl3)) #allows you to compare multiple models, ranks them
bictab(list(mdl1, mdl2, mdl3))
```



## Compare models using k-fold cross validation

```{r}
folds <- 10
fold_vec <- rep(1:folds, length.out = nrow(penguins_clean))

# want to randomize assignment of bins so that all light/heavy penguins
# aren't put in same test set.
# if you set seed, then it's easier to replicate
set.seed(42)
penguins_fold <- penguins_clean %>% 
  mutate(group = sample(fold_vec, size = n(), replace = FALSE))

table(penguins_fold$group)

test_df <- penguins_fold %>% 
  filter(group == 1)
train_df <- penguins_fold %>% 
  filter(group != 1)

```


Create function to calculate RMSE (root mean square error)
```{r}
# ex simple function
calc_mean <- function(x) {
  m <- sum(x) / length(x)
}
  
# real function we want to make for RMSE
calc_rmse <- function(x, y) {
  rmse <- (x - y)^2 %>% 
    mean() %>% 
    sqrt()
  return(rmse)  #give back to user what the final value is
}
```


```{r}
training_mdl1 <- lm(f1, data = train_df)
training_mdl2 <- lm(f2, data = train_df)
training_mdl3 <- lm(f3, data = train_df)

predict_test <- test_df %>% 
  mutate(model1 = predict(training_mdl1, test_df),
         model2 = predict(training_mdl2, test_df),
         model3 = predict(training_mdl3, test_df))

rmse_predict_test <- predict_test %>% 
  summarize(rmse_mdl1 = calc_rmse(model1, mass),
            rmse_mdl2 = calc_rmse(model2, mass),
            rmse_mdl3 = calc_rmse(model3, mass))
```


## Let's interate

```{r}
rmse_df <- data.frame()  #create empty df

for(i in 1:folds) {
  kfold_test_df <- penguins_fold %>% 
    filter(group == i)
  kfold_train_df <- penguins_fold %>% 
    filter(group != i)
  
  kfold_mdl1 <- lm(f1, data = kfold_train_df)
  kfold_mdl2 <- lm(f2, data = kfold_train_df)
  kfold_mdl3 <- lm(f3, data = kfold_train_df)
  
  kfold_pred_df <- kfold_test_df %>% 
    mutate(mdl1 = predict(kfold_mdl1, kfold_test_df),
           mdl2 = predict(kfold_mdl2, .),
           mdl3 = predict(kfold_mdl3, .)) #period is shorthand for putting in operating df

   kfold_rmse_df <- kfold_pred_df %>% 
    summarize(rmse_mdl1 = calc_rmse(mdl1, mass),
              rmse_mdl2 = calc_rmse(mdl2, mass),
              rmse_mdl3 = calc_rmse(mdl3, mass),
              test_gp = i)
  
   rmse_df <- bind_rows(rmse_df, kfold_rmse_df)
}


rmse_df %>% 
  summarize(mean_rmse_mdl1 = mean(rmse_mdl1),
            mean_rmse_mdl2 = mean(rmse_mdl2),
            mean_rmse_mdl3 = mean(rmse_mdl3))
```


# Finalize the model

```{r}
final_mdl <- lm(f2, data = penguins_clean)
```


Our final model:
`r equatiomatic::extract_eq(final_mdl, wrap = TRUE)`

And with coefficients:
`r equatiomatic::extract_eq(final_mdl, wrap = TRUE, use_coefs = TRUE)`












